Towards cross-middleware authentication and single sign-on for ARC Grid middleware

### 2009-02-01 doc/sec/papers/ISC09/ISC09.tex
Making pdflatex working

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2006/03/15
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallextended]{svjour3}     % onecolumn (second format)
\documentclass[twocolumn]{svjour3}         % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

@@

\section{Introduction} 
\label{sec:intro}
Grid provides the technology for wide-scale, cross-domain collaboration where users work with resources and share data which are distributed across administrative domains. This kind of collaboration is called Virtual Organization (VO)\cite{IFoster01}. The resources that contribute to the VO can come from independent administrative domains, and even require different security approaches to access control due to usage of different underlying middlewares. The same heterogeneous situation applies to users who need to access the VO resources, because users holding one kind of a credential may need to utilize multiple resources with different security approaches. To solve the above issues, Grid middlewares should provide the following abilities:
The user which participates a VO should be able to access data and resources without being required to constantly do authentication at each resource or service;
The VO should be able to easily manage the privileges and memberships of the users;
The VO together with its resources should be able to easily manage the authorization, given the condition that the resources are owned by different administrative domains.

In the current Grid community, the Grid Security Infrastructure (GSI)\cite{IFoster98} is the de facto standard for authentication and transport level security. It builds on the Public Key Infrastructure (PKI), an architecture based on the X.509 public key certificate. GSI implements some enhancement (such as certificate delegation) based on the standard SSL/TLS protocol. Mutual authentication is required by GSI, and is the default configuration for GSI based Grid deployment. The X.509 certificates are required for both the client and service sides in order to achieve mutual authentication. The X.509 certificates are issued by trusted third parties called certificate authorities (CA), and the CAs constitute trust federation which guarantees that two different X.509 certificates from different CAs can accomplish authentication against each other. Thus in the GSI based Grid system, if a user would like to access a Grid system, he/she should own a X.509 certificate which is issued by a CA that is trusted by other entities.

In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS)\cite{AlfieriR05} have partly solved them, to a different extent. The identity delegation and the X.509 proxy certificate initiated by GSI can achieve single sign-on, so that the need to re-enter the user’s pass phrase for authentication can be avoided by creating the proxy certificate.  As the most successful Grid authorization model, VOMS adopts the Attribute Certificate\cite{RFC3821link} approach. The Attribute Certificate is signed by the VOMS server and then bound to proxy certificates by the client.Attributes include the VO membership data which is associated with the proxy certificate’s identity. Eventually the resources can make access control decisions based on parsing the VO membership attributes from the Attribute Certificate.

The VO is essentially supposed to be able to provide resource sharing capability with resources running under heterogeneous systems/middlewares. This goal can be achieved using Web Service Architecture (WSA). This way, the Web Service technology has been utilized in Grid computing middlewares. The Web Service technology has been adopted by the Globus Toolkit\cite{GTlink}, gLite\cite{gLitelink}, as well as the Advanced Resource Connector (ARC)\cite{ARClink}. In fact, utilizing Web Service technology and providing seamless interoperability with other middlewares is one of the main goals of the new version of ARC middleware.

Unfortunately, the existing Grid security solutions such as GSI and VOMS do not apply to Web Service based applications very well due to the following reasons:
The GSS API based confidential communication has not been adopted by Web Service implementations.
The Attribute Certificate has not been adopted by any Web Service implementation at all.
Web Service implementations commonly use the Web Services Security (WS-Security)\cite{WSSeclink}, Security Assertion Markup Languages (SAML)\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
Finally, identity delegation is a good solution for user or user’s representative (acting on behalf of the user) to move/travel from one resource to another. But since the identity delegation has been completely coupled with the GSI confidential communication process, it is not possible for the existing identity delegation solution to be used by Web Service applications.

@@

\section{ARC Grid middleware}
\label{sec:arcmiddleware}
ARC (Advanced Resource Connector) is an open source Grid middleware solution released under Apache license. ARC middleware aims at developing self-organized, fault-tolerant, non-intrusive, easy-manageable Grid middleware\cite{MEllert07}. The current production version of ARC provides Grid services for job submission and management, resource characterization, resource aggregation and discovery, basic data management, integration of Grid security solution, and so on. This release has been deployed and used in production environment, and is one of the widely deployed Grid middlewares in Europe.

The next generation of ARC components is developed by the KnowARC project\cite{KnowARClink}, based on the functionality and capabilities of the classic ARC middleware. It aims at implementing a Web Service oriented Grid middleware which will provide higher levels of resource and user abstraction through well-defined Web Service interface\cite{KnowARCDesignlink} in order to provide interoperability with other service-based Grid middlewares, as well as other Web Service compatible applications.

@@

Dotted line in Figure \ref{fig:HED} shows an alternative path for the information to propagate among MCCs. A service administrator can configure the MCCs according to the interoperability requirements with the other part. For instance, the configuration marked with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP message mechanism (see WSE’s SoapSender and SoapReceiver)\cite{WSElink}. Another configuration could be SOAP over HTTPg (HTTP over GSI communication protocol) which is needed to interoperate with services like the Storage Resource Manager (SRM)\cite{AShoshani03} service. This shows the flexibility of HED in terms of protocols support.

\begin{figure}
\includegraphics[width=0.75\textwidth]{HED.eps}
\caption{The example of Host Environment Daemon deployed with A-REX and File services}
\label{fig:HED}
\end{figure}
HED contains a framework for implementing and enforcing authentication and authorization. Each MCC or Service has a common interface for implementing various authentication and authorization functionality. Each functionality can be implemented as a pluggable and configurable component (plug-ins) called SecHandler, which is implemented as a C++ class and provides a method for processing message that travels through MCC or Service. Each MCC or Service is usually configured with two queues of SecHandler – one for the incoming message and one for the outgoing message respectively. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well. In Figure \ref{fig:HED}, the “AuthZ” and “AuthN” sub-modules inside MCCs and Services are examples of SecHandler.

@@

\section{Cross-Middleware Authentication and Single Sign-On}
\label{sec:siglesignon}
We investigated the authentication and single sign-on issue in the scenario when a user or user’s representative (acting on behalf of this user) needs to access resources across different middlewares. Our work focuses on the following five major aspects:
Integration of community authentication – to utilize the standard-compatible community authentication mechanism from the Shibboleth\cite{Shiblink} Identity Provider software when users authenticate against the Grid services.
Short lived credentials service – to issue the short lived X.509 certificates for users authenticated using the community authentication, in order to interoperate with the services that depend on PKI based authentication.
Credential delegation service – to process the X.509 credential delegation as a normal Web Service, in order to provide single sign-on ability.
Message level authentication – to provide message level authentication by implementing WS-Security.
SAML token based delegation service – to process the delegation of an WS Security SAML token in order to provide single sign-on for SOAP message level authentication.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAML2SSO.eps}
\caption{SAML2.0 SSO profile in ARC}
\label{fig:SAML2SSO}
\end{figure}
Here is the explanation of steps shown in Figure \ref{fig:SAML2SSO}:
1.The client uses the user agent interface to launch a HTTP request including the IdP name (to which the user belongs) to the service side. The endpoint of the SP (Service Provider) service is the same as that of the other target services, except the last part of the endpoint is “saml2sp” which is specific for pointing to the SP Service. Note that we use Identity Provider (IdP) name here to simplify the IdP discovery process in order to avoid the IdP discovery process, because we suppose that the user who would access the target services should better know where is he from initially.
2.The SP Service searches the metadata (we use the same metadata format as defined in the Shibboleth) and gets the location of the single sign-on service (hosted in IdP) and also the location of assertion consuming service (hosted in this SP itself). Then SP Service uses obtained information to compose and issue the SAML authentication request also by utilizing own X.509 certificate (note that in the SAML2.0 SSO profile, X.509 certificates are still needed for IdP and SP) and sends the response back to the user agent.
3.User agent sends the SAML authentication request to the Identity Provider.
4.Identity Provider requires an act of authentication. The authentication mechanism is outside of the SAML2.0 SSO profile. The Shibboleth IdP implementation uses login handlers for authentication. The current user agent implementation is compatible with the Username/Password login handler of Shibboleth IdP. Through the HTTP protocol, the user agent will feed IdP with the username/password which has been given by the caller of user agent interface.
5.Once the authentication has been succeeded, the IdP issues a SAML response including SAML assertion encrypted by destination SP’s public key, and then this SAML response will be delivered by the user agent to the Service Provider.
6.The SP Service verifies and checks the SAML response, decrypts and stores the SAML assertion into session/connection context. That assertion includes the AuthnStatement and AttributeStatement elements.
7.The WS client launches the Grid/Web Service request via the same connection as the one which is used by user agent to contact SP Service.
8.The Grid/Web Service checks the AuthnStatement stored in the session context to see if the session is still valid. If valid, service handles the  request and returns the response to the WS client. Note that in the current implementation the service requires WS client to reuse same TCP connection as the one used by user agent is step 5, in order to guarantee  validity of the SSO result.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{SLCS.eps}
\caption{Short lived credential service}
\label{fig:SLCS}
\end{figure}
The SLCS service is also a Web Service implemented using ARC service interface. The SLCS client is implemented as command-line interface (CLI) utility which implements the user agent and WS client functionalities. The whole process of SLCS invocation is shown in Figure \ref{fig:SLCS} (steps 1‑8). The process resembles one shown in in Figure \ref{fig:SAML2SSO}, except that steps 7 and 8 use SLCS certificate request and response.

@@

It should be noticed that the process of composing the distinguished name (DN) for the certificate is a critical issue for the SLCS service. Since the Shibboleth Identity Provider uses the eduPerson schema\cite{eduSchemalink} for the composing the SAML Attribute we pick the relatively distinguishable attribute “eduPersonPrincipalName” for the DN. A typical example of the eduPersonPrincipalName  could be alice@example.org. In such case composed DN is for example “/O=knowarc/OU=example.org/CN=alice”.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{Delegation.eps}
\caption{Credential delegation service and client}
\label{fig:Delegation}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\textwidth]{Delegation_protocol.eps}
\caption{Credential delegation protocol}
\label{fig:Delegation_protocol}
\end{figure}
For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in Figure \ref{fig:Delegation}), should use the user’s X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential. For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in Figure \ref{fig:Delegation}), should also use the user’s X.509 credential or a delegated credential for authentication and secure communication.

@@

\subsection{Message Level Authentication}
\label{sec:msgauthn}
Transport level security (TLS or SSL) is not sufficient for securing Web Service, because the secure channel created by SSL/TLS makes it impossible to differentially protect the SOAP message, e.g. to encrypt or sign only particular components of the SOAP message, which is relevant when non-sensitive portions of the message need to be accessed or changed by intermediate actors. Additionally, SSL/TLS is incapable of providing end-to-end protection for SOAP message which might flow through multiple intermediate actors, because it only allows each hop to be protected with the resultant security gaps at intermediate actors. Web Services Security (WS-Security)\cite{WSSeclink} specification provides means for applying security to Web Service.

@@

The relying party is supposed to trust the issuing authority. When it receives a message from the asserting entity, it will check the SAML assertion based on its predetermined trust relationship with the SAML issuing authority, and check the signature of the SOAP message based on the public key in the SAML assertion without directly trust relationship with attesting entity (holder of this public key).
In terms of verification, since the proxy certificates have been introduced in IETF (RFC 3820)\cite{RFC3820link}, and OpenSSL has also implemented RFC3820, it is not difficult to use the proxy certificate for the X.509 Token profile to issue an X.509 token.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAMLDelegation.eps}
\caption{SAML Token delegation service and client}
\label{fig:SAMLDelegation}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAMLDelegationMessage.eps}
\caption{SOAP message header with SAML delegation token inside}
\label{fig:SAMLDelegationMessage}
\end{figure}
The delegation process is similar to the process in the X.509 credential delegation service, with the following differences:
1.Transport level communication between delegation client and service is not based on mutual authentication between the proxy certificate and the service certificate; instead, it is based on mutual authentication between the service certificates.
2.Delegation service will not generate any certificate request; instead, the delegation client will pick the public key from the peer certificate after a successful SSL/TLS authentication. So step 2 of the delegation protocol is just some confirmation information which informs delegation client to proceed.
3.For those multiple intermediate services through which a SOAP message flows, delegation service does not need to appear, instead, it only needs to appear on the end of the SOAP message flow.
4.SAML Token might include one or a few <saml:Attribute> attributes which carry the attribute information related to the subject, instead of the Attribute Certificate which is used in the VOMS proxy certificate for carrying the attribute information.

@@

\section{Related Work}
\label{sec:relatedwork}
Similar work that integrates Shibboleth and Grid middleware does exist. GridShib\cite{VWelch05},\cite{TBarton06} uses Shibboleth to get entity’s attribute for the authorization of services developed in Globus Tookit. Unlike the work described in this paper, the GridShib does not use Shibboleth’s native login handlers for authentication to Shibboleth IdP, instead, a new login handler specific for X.509 based authentication is developed for the Shibboleth IdP, and a credential service is developed for acting as an online-CA, authenticating the user, and issuing short-lived X.509 credential. Moreover, this paper provides a single sign-on solution for the user with community credentials authenticating to Grid services by directly using SAML2.0 SSO profile.

SWITCH (The Swiss Education and Research Network) short-lived credential service (SLCS)\cite{switchslcslink} provides a SLCS service and a client. The advantage of our work comparing to the SWTCH’s solution is that the SLCS service and client are based on Web Services, so that they can be implemented in any Web Service container while keeping interoperability.

In terms of delegation, apart from the credential delegation mechanism of GSI\cite{IFoster98},{VWelch04}, several authors\cite{MAhsant04} propose a delegation protocol based on the WS-Trust\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids. WS-Trust will also be adopted by the ARC middleware to express the specifications required to define delegation protocol in a standard way. Gridsite project implements a X.509 credential delegation solution\cite{GridSitelink} based on the Web Service, with which ARC delegation client can be made easily interoperable.

GSI plug-in for gSOAP\cite{GAloisio05} adds GSI support for gSOAP Web Service toolkit. The GSI MCC implemented in ARC provides samilar functionality. Moreover, the ARC solution is more flexible because GSI MCC of both service and client sides is pluggable, which means service and client developers of the ARC do not need to modify the service or client implementation in order to add or replace the GSI functionality.

@@

% For one-column wide figures use
\begin{figure}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics{example.eps}
% figure caption is below the figure
\caption{Please write your figure caption here}
\label{fig:1}       % Give a unique label
\end{figure}
%
% For two-column wide figures use
\begin{figure*}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[width=0.75\textwidth]{example.eps}
% figure caption is below the figure
\caption{Please write your figure caption here}
\label{fig:2}       % Give a unique label
\end{figure*}
%
% For tables use
\begin{table}
% table caption is above the table
\caption{Please write your table caption here}
\label{tab:1}       % Give a unique label
% For LaTeX tables use
\begin{tabular}{lll}
\hline\noalign{\smallskip}
first & second & third  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
number & number & number \\
number & number & number \\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

@@

\bibitem{KnowARCDesignlink}
Design document of new version ARC. https://www.knowarc.eu/documents/Knowarc_D1.1‑1_07.pdf

@@

\bibitem{VWelch05}
V. Welch, T. Barton, K. Keahey and F. Siebenlist. Attributes, Anonymity, and Access: Shibboleth and Globus Integration to Facilitate Grid Collaboration. 4th Annual PKI R&D Workshop, 2005.

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2006/03/15
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallextended]{svjour3}     % onecolumn (second format)
\documentclass[twocolumn]{svjour3}         % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

@@

\section{Introduction} 
\label{sec:intro}
Grid provides the technology for wide-scale, cross-domain collaboration where users work with resources and share data which are distributed across administrative domains. This kind of collaboration is called Virtual Organization (VO)~\cite{IFoster01}. The resources that contribute to the VO can come from independent administrative domains, and even require different security approaches to access control due to usage of different underlying middlewares. The same heterogeneous situation applies to users who need to access the VO resources, because users holding one kind of a credential may need to utilize multiple resources with different security approaches. To solve the above issues, Grid middlewares should provide the following abilities:
The user which participates a VO should be able to access data and resources without being required to constantly do authentication at each resource or service;
The VO should be able to easily manage the privileges and memberships of the users;
The VO together with its resources should be able to easily manage the authorization, given the condition that the resources are owned by different administrative domains.

In the current Grid community, the Grid Security Infrastructure (GSI)~\cite{IFoster98} is the de facto standard for authentication and transport level security. It builds on the Public Key Infrastructure (PKI), an architecture based on the X.509 public key certificate. GSI implements some enhancement (such as certificate delegation) based on the standard SSL/TLS protocol. Mutual authentication is required by GSI, and is the default configuration for GSI based Grid deployment. The X.509 certificates are required for both the client and service sides in order to achieve mutual authentication. The X.509 certificates are issued by trusted third parties called certificate authorities (CA), and the CAs constitute trust federation which guarantees that two different X.509 certificates from different CAs can accomplish authentication against each other. Thus in the GSI based Grid system, if a user would like to access a Grid system, he/she should own a X.509 certificate which is issued by a CA that is trusted by other entities.

In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS)~\cite{AlfieriR05} have partly solved them, to a different extent. The identity delegation and the X.509 proxy certificate initiated by GSI can achieve single sign-on, so that the need to re-enter the user’s pass phrase for authentication can be avoided by creating the proxy certificate.  As the most successful Grid authorization model, VOMS adopts the Attribute Certificate~\cite{RFC3821link} approach. The Attribute Certificate is signed by the VOMS server and then bound to proxy certificates by the client.Attributes include the VO membership data which is associated with the proxy certificate’s identity. Eventually the resources can make access control decisions based on parsing the VO membership attributes from the Attribute Certificate.

The VO is essentially supposed to be able to provide resource sharing capability with resources running under heterogeneous systems/middlewares. This goal can be achieved using Web Service Architecture (WSA). This way, the Web Service technology has been utilized in Grid computing middlewares. The Web Service technology has been adopted by the Globus Toolkit~\cite{GTlink}, gLite~\cite{gLitelink}, as well as the Advanced Resource Connector (ARC)~\cite{ARClink}. In fact, utilizing Web Service technology and providing seamless interoperability with other middlewares is one of the main goals of the new version of ARC middleware.

Unfortunately, the existing Grid security solutions such as GSI and VOMS do not apply to Web Service based applications very well due to the following reasons:
The GSS API based confidential communication has not been adopted by Web Service implementations.
The Attribute Certificate has not been adopted by any Web Service implementation at all.
Web Service implementations commonly use the Web Services Security (WS-Security)~\cite{WSSeclink}, Security Assertion Markup Languages (SAML)~\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
Finally, identity delegation is a good solution for user or user’s representative (acting on behalf of the user) to move/travel from one resource to another. But since the identity delegation has been completely coupled with the GSI confidential communication process, it is not possible for the existing identity delegation solution to be used by Web Service applications.

@@

\section{ARC Grid middleware}
\label{sec:arcmiddleware}
ARC (Advanced Resource Connector) is an open source Grid middleware solution released under Apache license. ARC middleware aims at developing self-organized, fault-tolerant, non-intrusive, easy-manageable Grid middleware~\cite{MEllert07}. The current production version of ARC provides Grid services for job submission and management, resource characterization, resource aggregation and discovery, basic data management, integration of Grid security solution, and so on. This release has been deployed and used in production environment, and is one of the widely deployed Grid middlewares in Europe.

The next generation of ARC components is developed by the KnowARC project~\cite{KnowARClink}, based on the functionality and capabilities of the classic ARC middleware. It aims at implementing a Web Service oriented Grid middleware which will provide higher levels of resource and user abstraction through well-defined Web Service interface~\cite{KnowARCDesignlink} in order to provide interoperability with other service-based Grid middlewares, as well as other Web Service compatible applications.

@@

Dotted line in Figure \ref{fig:HED} shows an alternative path for the information to propagate among MCCs. A service administrator can configure the MCCs according to the interoperability requirements with the other part. For instance, the configuration marked with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP message mechanism (see WSE’s SoapSender and SoapReceiver)~\cite{WSElink}. Another configuration could be SOAP over HTTPg (HTTP over GSI communication protocol) which is needed to interoperate with services like the Storage Resource Manager (SRM)~\cite{AShoshani03} service. This shows the flexibility of HED in terms of protocols support.

\begin{figure}
\includegraphics[width=0.75\textwidth]{HED.png}
\caption{The example of Host Environment Daemon deployed with A-REX and File services}
\label{fig:HED}
\end{figure}
HED contains a framework for implementing and enforcing authentication and authorization. Each MCC or Service has a common interface for implementing various authentication and authorization functionality. Each functionality can be implemented as a pluggable and configurable component (plug-ins) called SecHandler, which is implemented as a C++ class and provides a method for processing message that travels through MCC or Service. Each MCC or Service is usually configured with two queues of SecHandler – one for the incoming message and one for the outgoing message respectively. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well. In Figure \ref{fig:HED}, the “AuthZ” and “AuthN” sub-modules inside MCCs and Services are examples of SecHandler.

@@

\section{Cross-Middleware Authentication and Single Sign-On}
\label{sec:siglesignon}
We investigated the authentication and single sign-on issue in the scenario when a user or user’s representative (acting on behalf of this user) needs to access resources across different middlewares. Our work focuses on the following five major aspects:
Integration of community authentication – to utilize the standard-compatible community authentication mechanism from the Shibboleth~\cite{Shiblink} Identity Provider software when users authenticate against the Grid services.
Short lived credentials service – to issue the short lived X.509 certificates for users authenticated using the community authentication, in order to interoperate with the services that depend on PKI based authentication.
Credential delegation service – to process the X.509 credential delegation as a normal Web Service, in order to provide single sign-on ability.
Message level authentication – to provide message level authentication by implementing WS-Security.
SAML token based delegation service – to process the delegation of an WS Security SAML token in order to provide single sign-on for SOAP message level authentication.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAML2SSO.png}
\caption{SAML2.0 SSO profile in ARC}
\label{fig:SAML2SSO}
\end{figure}
Here is the explanation of steps shown in Figure \ref{fig:SAML2SSO}:
1.The client uses the user agent interface to launch a HTTP request including the IdP name (to which the user belongs) to the service side. The endpoint of the SP (Service Provider) service is the same as that of the other target services, except the last part of the endpoint is “saml2sp” which is specific for pointing to the SP Service. Note that we use Identity Provider (IdP) name here to simplify the IdP discovery process in order to avoid the IdP discovery process, because we suppose that the user who would access the target services should better know where is he from initially.
2.The SP Service searches the metadata (we use the same metadata format as defined in the Shibboleth) and gets the location of the single sign-on service (hosted in IdP) and also the location of assertion consuming service (hosted in this SP itself). Then SP Service uses obtained information to compose and issue the SAML authentication request also by utilizing own X.509 certificate (note that in the SAML2.0 SSO profile, X.509 certificates are still needed for IdP and SP) and sends the response back to the user agent.
3.User agent sends the SAML authentication request to the Identity Provider.
4.Identity Provider requires an act of authentication. The authentication mechanism is outside of the SAML2.0 SSO profile. The Shibboleth IdP implementation uses login handlers for authentication. The current user agent implementation is compatible with the Username/Password login handler of Shibboleth IdP. Through the HTTP protocol, the user agent will feed IdP with the username/password which has been given by the caller of user agent interface.
5.Once the authentication has been succeeded, the IdP issues a SAML response including SAML assertion encrypted by destination SP’s public key, and then this SAML response will be delivered by the user agent to the Service Provider.
6.The SP Service verifies and checks the SAML response, decrypts and stores the SAML assertion into session/connection context. That assertion includes the AuthnStatement and AttributeStatement elements.
7.The WS client launches the Grid/Web Service request via the same connection as the one which is used by user agent to contact SP Service.
8.The Grid/Web Service checks the AuthnStatement stored in the session context to see if the session is still valid. If valid, service handles the  request and returns the response to the WS client. Note that in the current implementation the service requires WS client to reuse same TCP connection as the one used by user agent is step 5, in order to guarantee  validity of the SSO result.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{SLCS.png}
\caption{Short lived credential service}
\label{fig:SLCS}
\end{figure}
The SLCS service is also a Web Service implemented using ARC service interface. The SLCS client is implemented as command-line interface (CLI) utility which implements the user agent and WS client functionalities. The whole process of SLCS invocation is shown in Figure \ref{fig:SLCS} (steps 1‑8). The process resembles one shown in in Figure \ref{fig:SAML2SSO}, except that steps 7 and 8 use SLCS certificate request and response.

@@

It should be noticed that the process of composing the distinguished name (DN) for the certificate is a critical issue for the SLCS service. Since the Shibboleth Identity Provider uses the eduPerson schema~\cite{eduSchemalink} for the composing the SAML Attribute we pick the relatively distinguishable attribute “eduPersonPrincipalName” for the DN. A typical example of the eduPersonPrincipalName  could be alice@example.org. In such case composed DN is for example “/O=knowarc/OU=example.org/CN=alice”.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{Delegation.png}
\caption{Credential delegation service and client}
\label{fig:Delegation}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\textwidth]{Delegation_protocol.png}
\caption{Credential delegation protocol}
\label{fig:Delegation_protocol}
\end{figure}
For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in Figure \ref{fig:Delegation}), should use the user’s X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential. For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in Figure \ref{fig:Delegation}), should also use the user’s X.509 credential or a delegated credential for authentication and secure communication.

@@

\subsection{Message Level Authentication}
\label{sec:msgauthn}
Transport level security (TLS or SSL) is not sufficient for securing Web Service, because the secure channel created by SSL/TLS makes it impossible to differentially protect the SOAP message, e.g. to encrypt or sign only particular components of the SOAP message, which is relevant when non-sensitive portions of the message need to be accessed or changed by intermediate actors. Additionally, SSL/TLS is incapable of providing end-to-end protection for SOAP message which might flow through multiple intermediate actors, because it only allows each hop to be protected with the resultant security gaps at intermediate actors. Web Services Security (WS-Security)~\cite{WSSeclink} specification provides means for applying security to Web Service.

@@

The relying party is supposed to trust the issuing authority. When it receives a message from the asserting entity, it will check the SAML assertion based on its predetermined trust relationship with the SAML issuing authority, and check the signature of the SOAP message based on the public key in the SAML assertion without directly trust relationship with attesting entity (holder of this public key).
In terms of verification, since the proxy certificates have been introduced in IETF (RFC 3820)~\cite{RFC3820link}, and OpenSSL has also implemented RFC3820, it is not difficult to use the proxy certificate for the X.509 Token profile to issue an X.509 token.

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAMLDelegation.png}
\caption{SAML Token delegation service and client}
\label{fig:SAMLDelegation}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAMLDelegationMessage.png}
\caption{SOAP message header with SAML delegation token inside}
\label{fig:SAMLDelegationMessage}
\end{figure}
The delegation process is similar to the process in the X.509 credential delegation service, with the following differences:
1.Transport level communication between delegation client and service is not based on mutual authentication between the proxy certificate and the service certificate; instead, it is based on mutual authentication between the service certificates.
2.Delegation service will not generate any certificate request; instead, the delegation client will pick the public key from the peer certificate after a successful SSL/TLS authentication. So step 2 of the delegation protocol is just some confirmation information which informs delegation client to proceed.
3.For those multiple intermediate services through which a SOAP message flows, delegation service does not need to appear, instead, it only needs to appear on the end of the SOAP message flow.
4.SAML Token might include one or a few <saml:Attribute> attributes which carry the attribute information related to the subject, instead of the Attribute Certificate which is used in the VOMS proxy certificate for carrying the attribute information.

@@

\section{Related Work}
\label{sec:relatedwork}
Similar work that integrates Shibboleth and Grid middleware does exist. GridShib~\cite{VWelch05},\cite{TBarton06} uses Shibboleth to get entity’s attribute for the authorization of services developed in Globus Tookit. Unlike the work described in this paper, the GridShib does not use Shibboleth’s native login handlers for authentication to Shibboleth IdP, instead, a new login handler specific for X.509 based authentication is developed for the Shibboleth IdP, and a credential service is developed for acting as an online-CA, authenticating the user, and issuing short-lived X.509 credential. Moreover, this paper provides a single sign-on solution for the user with community credentials authenticating to Grid services by directly using SAML2.0 SSO profile.

SWITCH (The Swiss Education and Research Network) short-lived credential service (SLCS)~\cite{switchslcslink} provides a SLCS service and a client. The advantage of our work comparing to the SWTCH’s solution is that the SLCS service and client are based on Web Services, so that they can be implemented in any Web Service container while keeping interoperability.

In terms of delegation, apart from the credential delegation mechanism of GSI~\cite{IFoster98},{VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol based on the WS-Trust~\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids. WS-Trust will also be adopted by the ARC middleware to express the specifications required to define delegation protocol in a standard way. Gridsite project implements a X.509 credential delegation solution~\cite{GridSitelink} based on the Web Service, with which ARC delegation client can be made easily interoperable.

GSI plug-in for gSOAP~\cite{GAloisio05} adds GSI support for gSOAP Web Service toolkit. The GSI MCC implemented in ARC provides samilar functionality. Moreover, the ARC solution is more flexible because GSI MCC of both service and client sides is pluggable, which means service and client developers of the ARC do not need to modify the service or client implementation in order to add or replace the GSI functionality.

@@

%% For one-column wide figures use
%\begin{figure}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:1}       % Give a unique label
%\end{figure}
%%
%% For two-column wide figures use
%\begin{figure*}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[width=0.75\textwidth]{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:2}       % Give a unique label
%\end{figure*}
%
%% For tables use
%\begin{table}
%% table caption is above the table
%\caption{Please write your table caption here}
%\label{tab:1}       % Give a unique label
%% For LaTeX tables use
%\begin{tabular}{lll}
%\hline\noalign{\smallskip}
%first & second & third  \\
%\noalign{\smallskip}\hline\noalign{\smallskip}
%number & number & number \\
%number & number & number \\
%\noalign{\smallskip}\hline
%\end{tabular}
%\end{table}

@@

\bibitem{KnowARCDesignlink}
Design document of new version ARC. https://www.knowarc.eu/documents/Knowarc\_D1.1-1\_07.pdf

@@

\bibitem{VWelch05}
V. Welch, T. Barton, K. Keahey and F. Siebenlist. Attributes, Anonymity, and Access: Shibboleth and Globus Integration to Facilitate Grid Collaboration. 4th Annual PKI R\&D Workshop, 2005.
